---
title: "Reproducible report example"
author: "Ivan C. Hanigan"
output:  pdf_document
---

# Some exploratory analysis
In this section we do some exploratory analysis of the NMMAPS data for
deaths in Chicago 1987-2000. The code, messages and intermediary
results are hidden in the resulting report document.

```{r, echo = FALSE, message = FALSE}
## We begin by reading in the data file:
## If using our own data we would use 'read.csv' or a similar tool to import data to R
# my.data <- read.csv('data/sampledata.csv',header=TRUE)
## for this example use data that are included in the dlnm package
library(dlnm)
# look at the structure of the data
# str(chicagoNMMAPS)
# summary(chicagoNMMAPS)
```

We made a simple scatter plot shown below

```{r, echo = FALSE, message = FALSE}
## make some plots. first by day
# with(chicagoNMMAPS, plot(date, cvd, type = "l"))
# we suspect a relationship between temperature and deaths
with(chicagoNMMAPS, plot(temp, cvd, pch = 16, cex = .6))
title(main = "A scatter plot of daily temperatures against deaths")
```

We ran some exploratory models. A Poisson GAM with smooth functions
on temperature and time was compared to a linear fit on temperature.

```{r, echo = FALSE, message = FALSE}
library(mgcv)
fit1 <- gam(cvd ~ s(temp) + s(time), data=chicagoNMMAPS, family = "poisson")
# we can access post-estimation summary statistics
# summary(fit1)
# or just plot the exposure-response function
plot(fit1, select = 1)
title(main = "The exposure-response function estimated using MGCV")
aic1 <- AIC(fit1)
# compare this to a linear term for temperature
aic0 <- AIC(gam(cvd ~ temp + s(time), data=chicagoNMMAPS, family = "poisson"))
# calculate the delta aic
aici <- aic1 - aic0
```

The result can be automatically inserted to the text. This model has
a delta AIC of `r round(aici,1)` (smoothed minus linear term).
